\section{Introduction}
\label{sec:introduction}

Digital Twins are increasingly presented as foundations for monitoring, simulation, prediction, optimisation, and decision support in manufacturing. Foundational and review literature emphasises virtual--physical linkage, lifecycle information, data integration, simulation, and synchronisation \cite{grieves2017digital,kritzinger2018digital,fuller2020digital,negri2017review,semeraro2021digital}. These perspectives are necessary, but they explain Digital Twin development mainly through architectures, data flows, and model integration. They say less about how the knowledge of operators is elicited, validated, and represented before formal modelling begins.

This omission matters for human-in-the-loop Digital Twins. Operators do not only use dashboards or respond to alarms. They interpret machine behaviour through situated cues such as sound, vibration, rhythm, surface finish, tool behaviour, alarm combinations, and deviations from expected response. Their actions often depend on context, uncertainty, escalation paths, production risk, and trust in the available evidence. Human-centred Digital Twin research has therefore begun to ask where humans fit in the Digital Twin loop and how collaboration, responsibility, and trust should be distributed between human actors and digital systems \cite{agrawal2023humans,longo2022ubiquitous,lee2004trust,parasuraman2000levels}. However, there remains a practical gap between recognising that operator knowledge matters and providing a repeatable method for turning situated operator explanations into model-ready knowledge.

We address that gap through the notion of an \emph{operator-strategy fragment}. An operator-strategy fragment is an intermediate knowledge object that relates a perceived cue, operating context, interpretation, possible action, rationale, uncertainty, risk, responsibility, and validation state. It is more structured than a raw field note, but less formal than an executable rule, SysML element, knowledge-graph structure, or domain-specific language construct. This intermediate status is important because premature formalisation can erase precisely the contextual judgement that makes operator knowledge useful.

The paper presents \emph{Notebook-to-OSL}, a hybrid elicitation-and-transformation method for early-stage Digital Twin development. The method is designed for apprenticeship-like situations where a researcher or engineer stands beside an operator and is trained in the work process. During the training session, the researcher captures low-friction digital field notes rather than forcing each observation into a schema. After the session, the notes are selected, annotated, clarified through cognitive-task-analysis-inspired prompts, validated with the operator, and transformed into model-ready candidates. OSL, short for Operator Strategy Language, is treated here as one possible downstream target representation. The contribution is not the complete OSL language; it is the traceable method that prepares situated operator knowledge for later formalisation.

The method is grounded in five bodies of literature. Contextual inquiry supports learning from real work in context and treating interpretation as a collaborative activity \cite{beyer1995apprenticing,beyer1997contextual}. Situated learning and cognitive apprenticeship justify the training situation as an elicitation setting in which expert reasoning becomes visible through demonstration, coaching, articulation, and reflection \cite{lave1991situated,collins1989cognitive}. Cognitive task analysis supports the focus on cues, difficult decisions, hidden judgments, and expert--novice differences \cite{crandall2006workingminds,klein1989critical,stanik2021acta}. Requirements engineering and systems engineering provide traceability, rationale, risk, source, status, and validation principles \cite{vanlamsweerde2009requirements,sebokStakeholder,sebokRequirements}. Finally, MBSE and semantic Digital Twin research motivate the need for model-ready fragments that can later feed SysML v2, requirements templates, knowledge graphs, or Digital Twin knowledge models \cite{omgSysml2024,karabulut2023ontologies,listl2024knowledge,braun2022graph}.

The paper addresses the following research questions:

\begin{description}[leftmargin=2.4cm,style=nextline]
    \item[RQ1] How can situated operator training be used as an elicitation setting for capturing operator strategies in Digital Twin development?
    \item[RQ2] How can low-friction digital field notes from CNC operator training be transformed into traceable, validated, model-ready operator-strategy fragments?
    \item[RQ3] Which annotation categories are needed to preserve context, cue interpretation, action logic, uncertainty, rationale, risk, responsibility, and validation needs during that transformation?
\end{description}

The paper makes four contributions. First, it positions apprenticeship-like operator training as a literature-grounded elicitation setting for human-in-the-loop Digital Twin development. Second, it proposes a hybrid workflow that combines contextual capture, post-session annotation, CTA-style clarification, operator validation, and model-ready transformation. Third, it introduces an annotation schema derived from contextual inquiry, cognitive task analysis, requirements engineering, and MBSE principles. Fourth, it provides a pilot-oriented evaluation protocol and an illustrative CNC example showing how a raw note can be transformed into a structured candidate for later formalisation.

The remainder of the paper is structured as follows. Section~\ref{sec:background} reviews the literature foundations and synthesises the gap. Section~\ref{sec:research-design} presents the research design, evaluation logic, and methodological positioning. Section~\ref{sec:method} describes the Notebook-to-OSL workflow. Section~\ref{sec:schema} defines and justifies the annotation schema. Section~\ref{sec:case} provides an illustrative CNC case. Section~\ref{sec:discussion} discusses implications, limitations, and evaluation needs. Section~\ref{sec:conclusion} concludes the paper.
