\section{Conclusion}
\label{sec:conclusion}

This paper proposed Notebook-to-OSL, a method for capturing operator strategies during situated CNC operator training and transforming raw laptop notes into structured candidate fragments for Digital Twin development. The method responds to a gap between informal operator practice and formal Digital Twin artefacts. Operators often use situated cues, contextual judgement, and experience-based reasoning that are difficult to obtain from machine data or documentation alone. If this knowledge is not captured carefully, Digital Twin development may lose the reasoning needed for meaningful human-in-the-loop support.

The proposed workflow separates raw note capture, post-session annotation, operator validation, and OSL candidate generation. This separation allows the researcher to remain present during training while still creating structured, traceable, and model-ready knowledge afterwards. The annotation schema preserves cues, conditions, interpretations, actions, rationale, exceptions, uncertainty, risks, responsibility, escalation paths, and validation needs. The illustrative CNC example showed how a note about cutting sound and tool wear can be transformed into a structured strategy fragment and an OSL candidate.

The next step is empirical application. A full study should apply the workflow during actual or simulated CNC training, analyse the resulting note transformations, validate fragments with operators, and evaluate whether the generated fragments support later OSL or Digital Twin modelling. If successful, the method can provide a practical bridge between apprenticeship-like learning in manufacturing and formal knowledge representations for Digital Twin systems.
