\documentclass[preprint,12pt]{elsarticle}

\usepackage[utf8]{inputenc}
\usepackage[T1]{fontenc}
\usepackage{lmodern}
\usepackage{booktabs}
\usepackage{tabularx}
\usepackage{array}
\usepackage{enumitem}
\usepackage{listings}
\usepackage{xcolor}
\usepackage{tikz}
\usepackage{hyperref}
\usetikzlibrary{arrows.meta,positioning,fit}

\hypersetup{
  colorlinks=true,
  linkcolor=black,
  citecolor=black,
  urlcolor=black
}

\lstdefinestyle{osl}{
  basicstyle=\ttfamily\small,
  breaklines=true,
  columns=fullflexible,
  frame=single,
  keepspaces=true
}

\journal{Advanced Engineering Informatics}

\begin{document}

\begin{frontmatter}

\title{From Situated Field Notes to Operator Strategy Models: Eliciting Manufacturing Knowledge for Digital Twin Development}

\author[hiost]{Martin Arthur Andersen\corref{cor1}}
\ead{nks.rap@gmail.com}
\cortext[cor1]{Corresponding author.}

\author[hiost]{Co-author Name}

\address[hiost]{Østfold University College, Halden, Norway}

\begin{abstract}
Digital Twin development in manufacturing depends on operational knowledge that is often unavailable in machine documentation, sensor streams, or formal process descriptions alone. Operators interpret sounds, vibrations, alarms, tooling behaviour, workarounds, and exceptions as part of situated practice, yet these interpretations are rarely transformed into traceable and model-ready artefacts before Digital Twin design begins. This paper presents Notebook-to-OSL, a hybrid elicitation-and-transformation method for capturing operator strategies during apprenticeship-like CNC operator training and preparing them for formalisation. The method combines contextual inquiry, situated learning, cognitive task analysis, requirements engineering, and model-based systems engineering. In the proposed workflow, a researcher captures low-friction digital field notes during training, then selects, annotates, clarifies, validates, and transforms them into operator-strategy fragments. The paper contributes a literature-grounded workflow, an annotation schema covering cues, context, interpretation, action, rationale, exception, uncertainty, risk, responsibility, traceability, and validation needs, and a pilot-oriented evaluation protocol for assessing whether fragments become model-ready candidates. The contribution is positioned as an intermediate knowledge layer between situated operator practice and formal Digital Twin artefacts such as requirements, decision-support rules, SysML elements, knowledge-graph structures, or future Operator Strategy Language constructs. The paper argues that human-in-the-loop Digital Twins require not only connected data and models, but also explicit methods for preserving the situated reasoning that makes operator support meaningful.
\end{abstract}

\begin{keyword}
Digital Twin \sep Operator strategy \sep Situated elicitation \sep Contextual inquiry \sep Cognitive task analysis \sep Requirements elicitation \sep Model-based systems engineering \sep Human-in-the-loop systems
\end{keyword}

\end{frontmatter}

\input{sections/01_introduction}
\input{sections/02_background}
\input{sections/03_research_design}
\input{sections/04_method}
\input{sections/05_annotation_schema}
\input{sections/06_illustrative_case}
\input{sections/07_discussion}
\input{sections/08_conclusion}

\section*{Declaration of generative AI and AI-assisted technologies in the manuscript preparation process}
During the preparation of this first manuscript draft, the authors used ChatGPT to support initial structuring, wording, and LaTeX drafting. The authors must review, revise, verify, and take full responsibility for the final content before submission.

\section*{CRediT authorship contribution statement}
To be completed before submission.

\section*{Declaration of competing interest}
The authors declare that they have no known competing financial interests or personal relationships that could have appeared to influence the work reported in this paper. This statement should be checked and revised before submission.

\section*{Funding}
Funding information should be added before submission. If no specific funding applies, use the journal's recommended no-funding statement.

\section*{Data availability}
No industrial data are included in this first draft. Any future empirical material should be anonymised and cleared with the industrial partner before publication.

\bibliographystyle{elsarticle-num}
\bibliography{references}

\end{document}
